---
title: "STAT472 Final"
author: "Katie Manzione"
date: "3/2/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(readr)
library(dplyr)
library(knitr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
```


```{r}
setwd("~/Desktop")
olympic <- read_csv("olympicdata.csv")
olympic<- olympic[,c("Sex", "Age", "Height", "Weight", "Team", "Games", "Year", "Sport", "Medal")]
olympic$Games <- gsub("\\d", "", olympic$Games)
olympic$Games <- trimws(olympic$Games)
olympic <- subset(olympic, Games == "Summer")
olympic$Medal[is.na(olympic$Medal)] <- 0
olympic <- olympic %>%
  mutate(Medal = recode(Medal, "Gold" = "1", "Silver" = "1", "Bronze" = "1"))
olympic$Medal <- as.numeric(olympic$Medal)
olympic <- na.omit(olympic)
olympic$Medal <- as.factor(olympic$Medal)
#olympic$Year <- as.factor(olympic$Year)

olympic <- olympic %>%
  mutate(Developed = Team) %>%
  mutate(Developed = recode(Developed, "Canada" = "1", "France"= "1", "Germany"= "1", "Italy"= "1", "Japan"= "1", "Great Britain"= "1", "United States"= "1", "Iceland"= "1", "Norway"= "1", "Switzerland"= "1", "Austria"= "1", "Belgium"= "1", "Denmark"= "1", "Finland"= "1", "Greece"= "1", "Ireland"= "1", "Luxembourg"= "1", "Netherlands"= "1", "Portugal"= "1", "Spain"= "1", "Sweden"= "1", "Australia" = "1", "Japan" = "1", "New Zealand" = "1", "Bulgaria" = "1", "Croatia"= "1", "Cyprus" = "1", "Estonia" = "1", "Czech Republic" = "1", "Hungary" = "1", "Latvia" = "1", "Lithuania" = "1", "Malta"  = "1", "Poland" = "1", "Romania" = "1", "Slovakia" = "1", "Slovenia" = "1"))
olympic$Developed[olympic$Developed != "1"] <- 0
head(olympic)
nrow(olympic)
```

```{r}
olympic$Sport <- as.factor(olympic$Sport)
```

--> use potentially: Cycling (7775), Diving (2011), Fencing (6537), Gymnastics (18271), Shooting (7260), Swimming (18776), Weightlifting (2974), Wrestling (5186)
--> there are "individual" sports. accounts for confounding of "how good the team is"


```{r}
sportList <- c("Cycling", "Diving", "Fencing", "Gymnastics", "Shooting", "Swimming", "Weightlifting", "Wrestling")


cycling <- subset(olympic, Sport == sportList[1])
cycling <- cycling[which(cycling$Year > 1960),]

diving <- subset(olympic, Sport == sportList[2])
diving <- diving[which(diving$Year > 1960),]

fencing <- subset(olympic, Sport == sportList[3])
fencing <- fencing[which(fencing$Year > 1960),]

gymnastics <- subset(olympic, Sport == sportList[4])
gymnastics <- gymnastics[which(gymnastics$Year > 1960),]

shooting <- subset(olympic, Sport == sportList[5])
shooting <- shooting[which(shooting$Year > 1960),]

swimming <- subset(olympic, Sport == sportList[6])
swimming <- swimming[which(swimming$Year > 1960),]

weightlifting <- subset(olympic, Sport == sportList[7])
weightlifting <- weightlifting[which(weightlifting$Year > 1960),]

wrestling <- subset(olympic, Sport == sportList[8])
wrestling <- wrestling[which(wrestling$Year > 1960),]


my_olympic <- rbind(cycling, diving, fencing, gymnastics, shooting, swimming, weightlifting, wrestling)
```

# EDA

## Developed

```{r, warning = FALSE}
olympic_developed1 <- ggplot(my_olympic) +
  geom_bar(aes(x=Year, fill=as.factor(Developed))) +
  ggtitle("Plot of Country Classification by Time")
olympic_developed1

olympic_developed2 <- ggplot(my_olympic) +
  geom_bar(aes(x=Medal, fill=as.factor(Developed)))+
  ggtitle("Barplot of Medal by Developed")
olympic_developed2
```


## Gender

```{r, warning=FALSE}
cycling_gender <- ggplot(cycling)+
  geom_bar(aes(x=Year, fill=as.factor(Sex))) +
  ggtitle("Cycling by Gender")

diving_gender <- ggplot(diving)+
  geom_bar(aes(x=Year, fill=as.factor(Sex))) +
  ggtitle("Diving by Gender")

fencing_gender <- ggplot(fencing)+
  geom_bar(aes(x=Year, fill=as.factor(Sex))) +
  ggtitle("Fencing by Gender")

gymnastics_gender <- ggplot(gymnastics)+
  geom_bar(aes(x=Year, fill=as.factor(Sex))) +
  ggtitle("Gymnastics by Gender")

shooting_gender <- ggplot(shooting)+
  geom_bar(aes(x=Year, fill=as.factor(Sex))) +
  ggtitle("Shooting by Gender")

swimming_gender <- ggplot(swimming)+
  geom_bar(aes(x=Year, fill=as.factor(Sex))) +
  ggtitle("Swimming by Gender")

weightlifting_gender <- ggplot(weightlifting)+
  geom_bar(aes(x=Year, fill=as.factor(Sex))) +
  ggtitle("Weightlifting by Gender")

wrestling_gender <- ggplot(wrestling)+
  geom_bar(aes(x=Year, fill=as.factor(Sex))) +
  ggtitle("Wrestling by Gender")
```

```{r}
ggarrange(cycling_gender, diving_gender, fencing_gender, gymnastics_gender, shooting_gender, swimming_gender, weightlifting_gender, wrestling_gender, ncol=2, nrow=2)
```

--subsetted to 1900 to present because 1900 was the first year that women competed at the olympics (olympics citation)

--cycling, fencing, weightlifting, wrestling looks like female proportions increase
--diving, gymnastics, shooting, swimming female/male proportions about equal


## Medals

```{r, warning=FALSE}
cycling_medal <- ggplot(cycling)+
  geom_bar(aes(x=Medal, fill=as.factor(Developed))) +
  ggtitle("Cycling Medals")

diving_medal <- ggplot(diving)+
  geom_bar(aes(x=Medal, fill=as.factor(Developed))) +
  ggtitle("Diving Medals")

fencing_medal <- ggplot(fencing)+
  geom_bar(aes(x=Medal, fill=as.factor(Developed))) +
  ggtitle("Fencing Medals")

gymnastics_medal <- ggplot(gymnastics)+
  geom_bar(aes(x=Medal, fill=as.factor(Developed))) +
  ggtitle("Gymnastics Medals")

shooting_medal <- ggplot(shooting)+
  geom_bar(aes(x=Medal, fill=as.factor(Developed))) +
  ggtitle("Shooting Medals")

swimming_medal <- ggplot(swimming)+
  geom_bar(aes(x=Medal, fill=as.factor(Developed))) +
  ggtitle("Swimming Medals")

weightlifting_medal <- ggplot(weightlifting)+
  geom_bar(aes(x=Medal, fill=as.factor(Developed))) +
  ggtitle("Weightlifting Medals")

wrestling_medal <- ggplot(wrestling)+
  geom_bar(aes(x=Medal, fill=as.factor(Developed))) +
  ggtitle("Wrestling Medals")
```


```{r}
ggarrange(cycling_medal, diving_medal, fencing_medal, gymnastics_medal, shooting_medal, swimming_medal, weightlifting_medal, wrestling_medal, ncol=2, nrow=2)
```

## Height

```{r, warning=FALSE}
cycling_height <- ggplot(cycling)+
  geom_density(aes(x=Height, col=as.factor(Medal))) +
  ggtitle("Cycling Height")

diving_height <- ggplot(diving)+
  geom_density(aes(x=Height, col=as.factor(Medal))) +
  ggtitle("Diving Height")

fencing_height <- ggplot(fencing)+
  geom_density(aes(x=Height, col=as.factor(Medal))) +
  ggtitle("Fencing Height")

gymnastics_height <- ggplot(gymnastics)+
  geom_density(aes(x=Height, col=as.factor(Medal))) +
  ggtitle("Gymnastics Height")

shooting_height <- ggplot(shooting)+
  geom_density(aes(x=Height, col=as.factor(Medal))) +
  ggtitle("Shooting Height")

swimming_height <- ggplot(swimming)+
  geom_density(aes(x=Height, col=as.factor(Medal))) +
  ggtitle("Swimming Height")

weightlifting_height <- ggplot(weightlifting)+
  geom_density(aes(x=Height, col=as.factor(Medal))) +
  ggtitle("Weightlifting Height")

wrestling_height <- ggplot(wrestling)+
  geom_density(aes(x=Height, col=as.factor(Medal))) +
  ggtitle("Wrestling Height")
```

```{r}
ggarrange(cycling_height, diving_height, fencing_height, gymnastics_height, shooting_height, swimming_height, weightlifting_height, wrestling_height, ncol=2, nrow=2)
```

--all relatively symmetric and normally distributed





```{r, warning=FALSE}
cycling_height_time <- ggplot(cycling)+
  geom_point(aes(x=Year, y = Height, col=as.factor(Medal))) +
  ggtitle("Cycling Height Over Time")

diving_height_time <- ggplot(diving)+
  geom_point(aes(x=Year,y=Height, col=as.factor(Medal))) +
  ggtitle("Diving Height Over Time")

fencing_height_time <- ggplot(fencing)+
  geom_point(aes(x=Year,y=Height, col=as.factor(Medal))) +
  ggtitle("Fencing Height Over Time")

gymnastics_height_time <- ggplot(gymnastics)+
  geom_point(aes(x=Year,y=Height, col=as.factor(Medal))) +
  ggtitle("Gymnastics Height Over Time")

shooting_height_time <- ggplot(shooting)+
  geom_point(aes(x=Year,y=Height, col=as.factor(Medal))) +
  ggtitle("Shooting Height Over Time")

swimming_height_time <- ggplot(swimming)+
  geom_point(aes(x=Year,y=Height, col=as.factor(Medal))) +
  ggtitle("Swimming Height Over Time")

weightlifting_height_time <- ggplot(weightlifting)+
  geom_point(aes(x=Year,y=Height, col=as.factor(Medal))) +
  ggtitle("Weightlifting Height Over Time")

wrestling_height_time <- ggplot(wrestling)+
  geom_point(aes(x=Year,y=Height, col=as.factor(Medal))) +
  ggtitle("Wrestling Height Over Time")
```

```{r}
ggarrange(cycling_height_time, diving_height_time, fencing_height_time, gymnastics_height_time, shooting_height_time, swimming_height_time, weightlifting_height_time, wrestling_height_time, ncol=2, nrow=2)
```

-- height distributions seem to increase in range over time, possibly due to greater number of athletes competing over time, in general.

# Weight

```{r, warning=FALSE}
cycling_weight <- ggplot(cycling)+
  geom_density(aes(x=Weight, col=as.factor(Medal))) +
  ggtitle("Cycling Weight")

diving_weight <- ggplot(diving)+
  geom_density(aes(x=Weight, col=as.factor(Medal))) +
  ggtitle("Diving Weight")

fencing_weight <- ggplot(fencing)+
  geom_density(aes(x=Weight, col=as.factor(Medal))) +
  ggtitle("Fencing Weight")

gymnastics_weight <- ggplot(gymnastics)+
  geom_density(aes(x=Weight, col=as.factor(Medal))) +
  ggtitle("Gymnastics Weight")

shooting_weight <- ggplot(shooting)+
  geom_density(aes(x=Weight, col=as.factor(Medal))) +
  ggtitle("Shooting Weight")

swimming_weight <- ggplot(swimming)+
  geom_density(aes(x=Weight, col=as.factor(Medal))) +
  ggtitle("Swimming Weight")

weightlifting_weight <- ggplot(weightlifting)+
  geom_density(aes(x=Weight, col=as.factor(Medal))) +
  ggtitle("Weightlifting Weight")

wrestling_weight <- ggplot(wrestling)+
  geom_density(aes(x=Weight, col=as.factor(Medal))) +
  ggtitle("Wrestling Weight")
```

```{r}
ggarrange(cycling_weight, diving_weight, fencing_weight, gymnastics_weight, shooting_weight, swimming_weight, weightlifting_weight, wrestling_weight, ncol=2, nrow=2)
```

--wrestling, weightlifting, shooting skewed right

--cycling, fencing, diving, gymnastics, swimming all relatively symmatric and normally distributed



```{r, warning=FALSE}
cycling_weight_time <- ggplot(cycling)+
  geom_point(aes(x=Year, y = Weight, col=as.factor(Medal))) +
  ggtitle("Cycling Weight Over Time")

diving_weight_time <- ggplot(diving)+
  geom_point(aes(x=Year,y=Weight, col=as.factor(Medal))) +
  ggtitle("Diving Weight Over Time")

fencing_weight_time <- ggplot(fencing)+
  geom_point(aes(x=Year,y=Weight, col=as.factor(Medal))) +
  ggtitle("Fencing Weight Over Time")

gymnastics_weight_time <- ggplot(gymnastics)+
  geom_point(aes(x=Year,y=Weight, col=as.factor(Medal))) +
  ggtitle("Gymnastics Weight Over Time")

shooting_weight_time <- ggplot(shooting)+
  geom_point(aes(x=Year,y=Weight, col=as.factor(Medal))) +
  ggtitle("Shooting Weight Over Time")

swimming_weight_time <- ggplot(swimming)+
  geom_point(aes(x=Year,y=Weight, col=as.factor(Medal))) +
  ggtitle("Swimming Weight Over Time")

weightlifting_weight_time <- ggplot(weightlifting)+
  geom_point(aes(x=Year,y=Weight, col=as.factor(Medal))) +
  ggtitle("Weightlifting Weight Over Time")

wrestling_weight_time <- ggplot(wrestling)+
  geom_point(aes(x=Year,y=Weight, col=as.factor(Medal))) +
  ggtitle("Wrestling Weight Over Time")
```

```{r}
ggarrange(cycling_weight_time, diving_weight_time, fencing_weight_time, gymnastics_weight_time, shooting_weight_time, swimming_weight_time, weightlifting_weight_time, wrestling_weight_time, ncol=2, nrow=2)
```

-- height distributions seem to increase in range over time, possibly due to greater number of athletes competing over time, in general.


# Modeling

## cycling

```{r}
cycling_glm <- glm(Medal ~ Age + Height + Weight + Developed, data=cycling, family = "binomial")
summary(cycling_glm)
```


prob of medal if from developed country is 66.33%

The coefficient for math says that, holding all other predictors constant, we will see 97.03% increase in the odds of getting a medal for athletes from developed countries, exp(0.678207) = 1.97034.

## diving

```{r}
diving_glm <- glm(Medal ~ Age + Height + Weight + Developed, data=diving, family = "binomial")
summary(diving_glm)
```

prob of medal if from developed country is 42.27%

The coefficient for math says that, holding all other predictors constant, we will see 26.78% decrease in the odds of getting a medal for athletes from developed countries, exp(-0.31172) = 0.732187.

## fencing

```{r}
fencing_glm <- glm(Medal ~ Age + Height + Weight + Developed, data=fencing, family = "binomial")
summary(fencing_glm)
```

prob of medal if from developed country is 55.10%

The coefficient for math says that, holding all other predictors constant, we will see 22.74% increase in the odds of getting a medal for athletes from developed countries, exp(.204904) = 1.2274.

## gymnastics

```{r}
gymnastics_glm <- glm(Medal ~ Age + Height + Weight + Developed, data=gymnastics, family = "binomial")
summary(gymnastics_glm)
```

prob of medal if from developed country is 30.38%

The coefficient for math says that, holding all other predictors constant, we will see 56.37% decrease in the odds of getting a medal for athletes from developed countries, exp(-0.829361) = 0.4363.

## shooting

```{r}
shooting_glm <- glm(Medal ~ Age + Height + Weight + Developed, data=shooting, family = "binomial")
summary(shooting_glm)
```


## swimming

```{r}
swimming_glm <- glm(Medal ~ Age + Height + Weight + Developed, data=swimming, family = "binomial")
summary(swimming_glm)
```

prob of medal if from developed country is 70.66%

The coefficient for math says that, holding all other predictors constant, we will see ~141% increase in the odds of getting a medal for athletes from developed countries, exp(0.879051) = 2.40861.


## weightlifting

```{r}
weightlifting_glm <- glm(Medal ~ Age + Height + Weight + Developed, data=weightlifting, family = "binomial")
summary(weightlifting_glm)
```

## wrestling

```{r}
wrestling_glm <- glm(Medal ~ Age + Height + Weight + Developed, data=wrestling, family = "binomial")
summary(wrestling_glm)
```


```{r}
round(exp(coef(cycling_glm)),3)
round(exp(coef(diving_glm)),3)
round(exp(coef(fencing_glm)),3)
round(exp(coef(gymnastics_glm)),3)
round(exp(coef(shooting_glm)),3)
round(exp(coef(swimming_glm)),3)
round(exp(coef(weightlifting_glm)),3)
round(exp(coef(wrestling_glm)),3)
```

--Indication of Developed Country is significant for cycling, fencing, diving, swimming, and gymnastics

# LASSO


library(glmnet)

olympic_LASSO_dat <- subset(olympic, select = c(Age, Sex, Year, Height, Weight, Developed, Medal))

x_vars <- model.matrix(Medal~. , olympic_LASSO_dat)[,-1]
y_var <- olympic_LASSO_dat$Medal
lambda_seq <- 10^seq(2, -2, by = -.1)

set.seed(472)
train = sample(1:nrow(x_vars), nrow(x_vars)/2)
x_test = (-train)
y_test = y_var[x_test]
 
cv_output <- cv.glmnet(x_vars[train,], y_var[train],
                       alpha = 1, lambda = lambda_seq, 
                       nfolds = 5, family = "binomial")
best_lambda <- cv_output$lambda.min

olympic_lasso <- glmnet(x_vars[train,], y_var[train], alpha = 1, lambda = best_lambda, family = "binomial")
pred_lasso <- predict(olympic_lasso, s = best_lambda, newx = x_vars[x_test,])

coef(olympic_lasso)


# Two-Sample Tests for equality of proportions

## Proportions of developed/not developed

```{r, echo = FALSE}
cycling_prop <- prop.test(x = c(length(which(cycling$Developed == 0)),length(which(cycling$Developed == 1))), n = c(length(cycling$Developed),length(cycling$Developed)), alternative = "two.sided")
cycling_prop
```

```{r, echo = FALSE}
diving_prop <- prop.test(x = c(length(which(diving$Developed == 0)),length(which(diving$Developed == 1))), n = c(length(diving$Developed),length(diving$Developed)), alternative = "two.sided")
diving_prop
```

```{r, echo = FALSE}
fencing_prop <- prop.test(x = c(length(which(fencing$Developed == 0)),length(which(fencing$Developed == 1))), n = c(length(fencing$Developed),length(fencing$Developed)), alternative = "two.sided")
fencing_prop
```

```{r, echo = FALSE}
gymnastics_prop <- prop.test(x = c(length(which(gymnastics$Developed == 0)),length(which(gymnastics$Developed == 1))), n = c(length(gymnastics$Developed),length(gymnastics$Developed)), alternative = "two.sided")
gymnastics_prop
```

```{r, echo = FALSE}
shooting_prop <- prop.test(x = c(length(which(shooting$Developed == 0)),length(which(shooting$Developed == 1))), n = c(length(shooting$Developed),length(shooting$Developed)), alternative = "two.sided")
shooting_prop
```

```{r, echo = FALSE}
swimming_prop <- prop.test(x = c(length(which(swimming$Developed == 0)),length(which(swimming$Developed == 1))), n = c(length(swimming$Developed),length(swimming$Developed)), alternative = "two.sided")
swimming_prop
```

```{r, echo = FALSE}
weightlifting_prop <- prop.test(x = c(length(which(weightlifting$Developed == 0)),length(which(weightlifting$Developed == 1))), n = c(length(weightlifting$Developed),length(weightlifting$Developed)), alternative = "two.sided")
weightlifting_prop
```

```{r, echo = FALSE}
wrestling_prop <- prop.test(x = c(length(which(wrestling$Developed == 0)),length(which(wrestling$Developed == 1))), n = c(length(wrestling$Developed),length(wrestling$Developed)), alternative = "two.sided")
wrestling_prop
```

## Of those who medalled, proportions of developed or not

```{r}
cyclingMedal <- subset(cycling, Medal == 1)
divingMedal <- subset(diving, Medal == 1)
fencingMedal <- subset(fencing, Medal == 1)
gymnasticsMedal <- subset(gymnastics, Medal == 1)
shootingMedal <- subset(shooting, Medal == 1)
swimmingMedal <- subset(swimming, Medal == 1)
weightliftingMedal <- subset(weightlifting, Medal == 1)
wrestlingMedal <- subset(wrestling, Medal == 1)
```


```{r, echo = FALSE}
cycling_propMedal <- prop.test(x = c(length(which(cyclingMedal$Developed == 0)),length(which(cyclingMedal$Developed == 1))), n = c(length(cyclingMedal$Developed),length(cyclingMedal$Developed)), alternative = "two.sided")
cycling_propMedal
```

```{r, echo = FALSE}
diving_propMedal <- prop.test(x = c(length(which(divingMedal$Developed == 0)),length(which(divingMedal$Developed == 1))), n = c(length(divingMedal$Developed),length(divingMedal$Developed)), alternative = "two.sided")
diving_propMedal
```

```{r, echo = FALSE}
fencing_propMedal <- prop.test(x = c(length(which(fencingMedal$Developed == 0)),length(which(fencingMedal$Developed == 1))), n = c(length(fencingMedal$Developed),length(fencingMedal$Developed)), alternative = "two.sided")
fencing_propMedal
```

```{r, echo = FALSE}
gymnastics_propMedal <- prop.test(x = c(length(which(gymnasticsMedal$Developed == 0)),length(which(gymnasticsMedal$Developed == 1))), n = c(length(gymnasticsMedal$Developed),length(gymnasticsMedal$Developed)), alternative = "two.sided")
gymnastics_propMedal
```

```{r, echo = FALSE}
shooting_propMedal <- prop.test(x = c(length(which(shootingMedal$Developed == 0)),length(which(shootingMedal$Developed == 1))), n = c(length(shootingMedal$Developed),length(shootingMedal$Developed)), alternative = "two.sided")
shooting_propMedal
```

```{r, echo = FALSE}
swimming_propMedal <- prop.test(x = c(length(which(swimmingMedal$Developed == 0)),length(which(swimmingMedal$Developed == 1))), n = c(length(swimmingMedal$Developed),length(swimmingMedal$Developed)), alternative = "two.sided")
swimming_propMedal
```

```{r, echo = FALSE}
weightlifting_propMedal <- prop.test(x = c(length(which(weightliftingMedal$Developed == 0)),length(which(weightliftingMedal$Developed == 1))), n = c(length(weightliftingMedal$Developed),length(weightliftingMedal$Developed)), alternative = "two.sided")
weightlifting_propMedal
```

```{r, echo = FALSE}
wrestling_propMedal <- prop.test(x = c(length(which(wrestlingMedal$Developed == 0)),length(which(wrestlingMedal$Developed == 1))), n = c(length(wrestlingMedal$Developed),length(wrestlingMedal$Developed)), alternative = "two.sided")
wrestling_propMedal
```

```{r, include = FALSE}
library(tidyverse) # data manip
library(ISLR) # data
library(GGally) # pairs plots
library(e1071) #svm
```
cycling, diving, fencing, gymnastics, shooting, swimming, weightlifting, wrestling

```{r}
nrow(diving)
0.7*nrow(diving)
0.3*nrow(diving)
diving_train <- cycling[1:1304,]
diving_test <- cycling[1305:1862,]
```

```{r}
0.7*nrow(fencing)
nrow(fencing)
nrow(fencing) - 3867
fencing_train <- fencing[1:3867,]
fencing_test <- fencing[3868:5523,]
```

```{r}
0.7*nrow(gymnastics)
0.3*nrow(gymnastics)
nrow(gymnastics) 
gymnastics_train <- gymnastics[1:10981,]
gymnastics_test <- gymnastics[10982:15686,]
```

```{r}
0.7*nrow(shooting)
0.3*nrow(shooting)
nrow(shooting)
shooting_train <- shooting[1:4665,]
shooting_test <- shooting[4666:6663,]
```

```{r}
0.7*nrow(swimming)
0.3*nrow(swimming)

swimming_train <- swimming[1:12520,]
swimming_test <- swimming[12521:17885,]
```

```{r}
0.7*nrow(weightlifting)
0.3*nrow(weightlifting)

weightlifting_train <- weightlifting[1:1892,]
weightlifting_test <- weightlifting[1893:2702,]
```

```{r}
0.7*nrow(wrestling)
0.3*nrow(wrestling)

wrestling_train <- wrestling[1:3253,]
wrestling_test <- wrestling[3254:4647,]
```



```{r}
cplot <- ggplot(cycling) +
  geom_point(aes(x=Weight, y=Height, col = as.factor(Medal)))
cplot

0.7*nrow(cycling)
0.3*nrow(cycling)

cycling_train <- cycling[1:5026,]
cycling_test <- cycling[5027:7181,]

7181 - 5026

log.reg.model_cyc <- glm(Medal ~ Height +Weight, data = cycling_train, family = "binomial")
summary(log.reg.model_cyc)

#get the estimated y value (0/1) for each point in the grid
yhat_cyc <- predict(log.reg.model_cyc, newdata = cycling_test, type = "response")
yhat_cyc <- (yhat_cyc > 0.5)*1
cycling_test$yhat <- yhat_cyc
summary(yhat_cyc)

#confMat_log <- table(actual=cycling_test$Medal, predicted=cycling_test$yhat)
#(confMat_log[1,2] + confMat_log[2,1])/dim(cycling_test)[1]



#cyc_fig <- ggplot()+ geom_tile(aes(x = Weight, y = Height, fill = as.factor(yhat)), data = cycling_test, alpha = .5) + labs(fill = "Predicted Y") + geom_point(data = cycling, aes(x = Weight, y = Height, col = as.factor(Medal))) + labs(col = "Observed Y")
#cyc_fig 


svmOut <-svm(Medal ~ Height , data = cycling_train, kernel = "linear", degree = 2)
summary(svmOut)
yhat_grid3 <- predict(svmOut, newdata = cycling_test, type = "response")
#yhat_grid3 <- (yhat_grid3 > 0.5)*1
cycling_test$yhat3 <- yhat_grid3
summary(yhat_grid3)

#confMat_svmLIN <- table(actual=cycling_test$Medal, predicted=cycling_test$yhat)
#(confMat_svmLIN[1,2] + confMat_svmLIN[2,1])/dim(cycling_test)[1]
#(confMat_svmLIN[1,1])/(confMat_svmLIN[1,1]+confMat_svmLIN[2,1])
#(confMat_svmLIN[1,1])/(confMat_svmLIN[1,1]+confMat_svmLIN[1,2])
#(confMat_svmLIN[2,2])/(confMat_svmLIN[2,2]+confMat_svmLIN[2,1])

#cyc_fig2 <- ggplot()+ geom_tile(aes(x = Weight, y = Height, fill = as.factor(yhat3)), data = cycling_test, alpha = .5) + labs(fill = "Predicted Y") + geom_point(data = cycling, aes(x = Weight, y = Height, col = as.factor(Medal))) + labs(col = "Observed Y")
#cyc_fig2


#work around: over-sampling-> sample more of the unbalanced class
#resample to 1000...on avg. 50/50 split to train
#canvas-->unbalanced classification issue slide (diff weight to fit the model...) -->adding weight to minor category


wt =ifelse(cycling_train$Medal==0, 1, 1000)
fit_logit_wt_cyc <-glm(Medal ~ Height + Weight, data = cycling_train,family=binomial,weights = wt)
#Log-odd ratio
pred_logit_wt_cyc<-predict(fit_logit_wt_cyc,newdata=cycling_test)

# Classify a sample as 0 when its predicted log-odd ratio is negative
y_hat_wt=ifelse(pred_logit_wt_cyc>=0, 1, 1000)
summary(y_hat_wt)

mean(y_hat_wt!=cycling_test$Medal)

cycling_test$yhat_wt <- y_hat_wt


cyc_fig_wt <- ggplot()+ geom_tile(aes(x = Weight, y = Height, fill = as.factor(yhat_wt)), data = cycling_test, alpha = .5) + labs(fill = "Predicted Y") + geom_point(data = cycling, aes(x = Weight, y = Height, col = as.factor(Medal))) + labs(col = "Observed Y")
cyc_fig_wt



svmOut_wt <-svm(Medal ~ Height , data = cycling_train, kernel = "linear", degree = 2, weight = wt)
summary(svmOut_wt)
yhat_grid_svmwt <- predict(svmOut_wt, newdata = cycling_test)
summary(yhat_grid_svmwt)
y_hat_wt=ifelse(pred_logit_wt_cyc>=0, 1, 1000)

#yhat_grid3 <- (yhat_grid3 > 0.5)*1
cycling_test$yhat3 <- yhat_grid3
summary(yhat_grid3)

cyc_fig2 <- ggplot()+ geom_tile(aes(x = Weight, y = Height, fill = as.factor(yhat3)), data = cycling_test, alpha = .5) + labs(fill = "Predicted Y") + geom_point(data = cycling, aes(x = Weight, y = Height, col = as.factor(Medal))) + labs(col = "Observed Y")
cyc_fig2
```


```{r}
library(rpart)
library(ROSE)
```

```{r}
#imb <- rpart(Medal ~ Height + Weight, data = cycling_train)
#pred.treeimb <- predict(imb, newdata = cycling_test)
#accuracy.meas(cycling_test$Medal, pred.treeimb[,2])

#data_balanced_over <- ovun.sample(Medal ~ Height + Weight, data = cycling_train, method = "over",N = nrow(cycling))$data
#table(data_balanced_over$Medal)

data.rose1 <- ROSE(Medal ~ Height + Weight , data = cycling_train, seed = 1)$data
table(data.rose1$Medal)

#data.rose2 <- ROSE(Medal ~ Height + Weight, data = cycling_test, seed = 1)$data
#table(data.rose2$Medal)

#cycling_rose <- rbind(data.rose1, data.rose2[,1:3])

#Precision = TP / (TP + FP) out of observations labeled as positive, how many are actually labeled positive
#Sensitivity = TP / (TP + FN) how many observations of positive class are labeled correctly.
#Specificity = TN / (TN + FP) how many observations of negative class are labeled correctly.


# Bayes Classifier

log.reg.model_cyc.rose <- glm(Medal ~ Height +Weight, data = data.rose1, family = "binomial")
#summary(log.reg.model_cyc.rose)

#get the estimated y value (0/1) for each point in the grid
yhat_cyc.logRose <- predict(log.reg.model_cyc.rose, newdata = cycling_test, type = "response")
yhat_cyc.logRose <- (yhat_cyc.logRose > 0.5)*1
cycling_test$yhat.logRose <- as.factor(yhat_cyc.logRose)
summary(cycling_test$yhat.logRose)

confMat_logRose <- table(actual=cycling_test$Medal, predicted=cycling_test$yhat.logRose)
(confMat_logRose[1,2] + confMat_logRose[2,1])/dim(cycling_test)[1]
(confMat_logRose[1,1])/(confMat_logRose[1,1]+confMat_logRose[2,1])
(confMat_logRose[1,1])/(confMat_logRose[1,1]+confMat_logRose[1,2])
(confMat_logRose[2,2])/(confMat_logRose[2,2]+confMat_logRose[2,1])
#error rate is 0.361949
#precision is 0.9173913
#sensitivity is 0.6552795
#specificity is 0.4887892



# Linear Rose SVM

svmOut_rose <-svm(Medal ~ Height +Weight, data = data.rose1, kernel = "linear", degree = 2)
summary(svmOut_rose)
yhat_grid_rosesvm <- predict(svmOut_rose, newdata = cycling_test, type = "response")
#yhat_grid3 <- (yhat_grid3 > 0.5)*1
cycling_test$yhat_rosesvm <- yhat_grid_rosesvm
summary(yhat_grid_rosesvm)

    
confMat_roseLINsvm<-table(actual=cycling_test$Medal, predicted=cycling_test$yhat_rosesvm)
(confMat_roseLINsvm[1,2] + confMat_roseLINsvm[2,1])/dim(cycling_test)[1]
(confMat_roseLINsvm[1,1])/(confMat_roseLINsvm[1,1]+confMat_roseLINsvm[2,1])
(confMat_roseLINsvm[1,1])/(confMat_roseLINsvm[1,1]+confMat_roseLINsvm[1,2])
(confMat_roseLINsvm[2,2])/(confMat_roseLINsvm[2,2]+confMat_roseLINsvm[2,1])
#error rate is 0.4566125
#precision is 0.9058219
#sensitivity is 0.547619
#specificity is 0.5067265



# Polynomial Rose SVM

svmOut_rose2 <-svm(Medal ~ Height +Weight, data = data.rose1, kernel = "polynomial", degree = 2)
summary(svmOut_rose2)
yhat_grid_rosesvm2 <- predict(svmOut_rose2, newdata = cycling_test, type = "response")
#yhat_grid3 <- (yhat_grid3 > 0.5)*1
cycling_test$yhat_rosesvm2 <- yhat_grid_rosesvm2
summary(yhat_grid_rosesvm2)


confMat_rosePOLYsvm<-table(cycling_test$Medal, cycling_test$yhat_rosesvm2)
(confMat_rosePOLYsvm[1,2] + confMat_rosePOLYsvm[2,1])/dim(cycling_test)[1]
(confMat_rosePOLYsvm[1,1])/(confMat_rosePOLYsvm[1,1]+confMat_rosePOLYsvm[2,1])
(confMat_rosePOLYsvm[1,1])/(confMat_rosePOLYsvm[1,1]+confMat_rosePOLYsvm[1,2])
(confMat_rosePOLYsvm[2,2])/(confMat_rosePOLYsvm[2,2]+confMat_rosePOLYsvm[2,1])
#error rate is 0.1308585
#precision is 0.8962536
#sensitivity is 0.9658385
#specificity is 0.03139013



# Polynomial Rose SVM

svmOut_rose3 <-svm(Medal ~ Height +Weight, data = data.rose1, kernel = "radial", degree = 2)
summary(svmOut_rose3)
yhat_grid_rosesvm3 <- predict(svmOut_rose3, newdata = cycling_test, type = "response")
#yhat_grid3 <- (yhat_grid3 > 0.5)*1
cycling_test$yhat_rosesvm3 <- yhat_grid_rosesvm3
summary(yhat_grid_rosesvm3)


confMat_roseRADsvm<-table(cycling_test$Medal, cycling_test$yhat_rosesvm3)
(confMat_roseRADsvm[1,2] + confMat_roseRADsvm[2,1])/dim(cycling_test)[1]
(confMat_roseRADsvm[1,1])/(confMat_roseRADsvm[1,1]+confMat_roseRADsvm[2,1])
(confMat_roseRADsvm[1,1])/(confMat_roseRADsvm[1,1]+confMat_roseRADsvm[1,2])
(confMat_roseRADsvm[2,2])/(confMat_roseRADsvm[2,2]+confMat_roseRADsvm[2,1])
#error rate is 0.3466357
#precision is 0.9134682
#sensitivity is 0.6775362
#specificity is 0.4439462



#contingency table (confusion matrix) --> actual medal/predicted medal

#lmitation -> two groups are not linearly separted 

#cyc_fig_rosesvm <- ggplot()+ geom_tile(aes(x = Weight, y = Height, fill = as.factor(yhat_rosesvm)), data = cycling_test, alpha = .5) + labs(fill = "Predicted Y") + geom_point(data = cycling, aes(x = Weight, y = Height, col = as.factor(Medal))) + labs(col = "Observed Y")
#cyc_fig_rosesvm


#gfig2 <- ggplot()+ geom_tile(aes(x = x1, y = x2, fill = as.factor(yhat2)), data = xgrid, alpha = .5) + labs(fill = "Predicted Y") + geom_point(data = mydf, aes(x = x1, y = x2, col = as.factor(y))) + labs(col = "Observed Y")
#gfig2 
```
```

